Graphical Tests for Power Comparison of Competing Designs

### 2012-03-30 intro.tex
tweaked power explanation

Lineups provide a way to statistically quantify the significance of the finding
\citet{majumder:2011}. Think of the linup as a set of $m$ plots, one of which is the real data. For one lineup and one observer, the significance level, $\alpha$ is $1/m$. This is the probability that the observer picks the plot of the real data by chance. It is the probability of picking the real data plot, when it is really *not* different from the other null plots, that is a Type I error. (In Figure \ref{lineup}, $m=20$.) The power of a test is the probability that the real data plot is identified when it really is different from the null plots. Power is usually difficult to calculate because it is requires specific use of alternative hypothesis to calculate the probability.  Work in \citet{niladri:2012} addresses this to some extent with measures on the quality of a lineup, how different numerically, as best it can be calculated, the real data plot is from the null plots. 
%
When there are multiple observers the significance level also depends on the number of independent observers ($n$). Let $n$ be the number of independent observers and $x_i$ the number of observers who picked plot $i$, $i \in \{1, ..., m\}$, from the lineup. Then $(x_1, x_2, ..., x_m)$ follows a multinomial distribution Mult$_{\pi_1, \pi_2, ..., \pi_m}$ with $\sum_i \pi_i = 1$, where $\pi_i$ is the probability that plot $i$ is picked by an observer, which we can estimate as $\widehat{\pi_i} = x_i/n$.

>>>>

Lineups provide a way to statistically quantify the significance of the finding
\citet{majumder:2011}. Think of the linup as a set of $m$ plots, one of which is the real data. For one lineup and one observer, the significance level, $\alpha$ is $1/m$. This is the probability that the observer picks the plot of the real data by chance. It is the probability of picking the real data plot, when it is really *not* different from the other null plots, that is, a Type I error. (In Figure \ref{lineup}, $m=20$.) The power of a test is the probability that the real data plot is identified when it really is different from the null plots. Power is usually difficult to calculate because it is requires specific use of alternative hypothesis to calculate the probability.  Work in \citet{niladri:2012} addresses this to some extent with measures on the quality of a lineup, how different numerically, as best it can be calculated, the real data plot is from the null plots. 
%

When there are multiple observers the significance level also depends on the number of independent observers ($n$). Let $n$ be the number of independent observers and $x_i$ the number of observers who picked plot $i$, $i \in \{1, ..., m\}$, from the lineup. Then $(x_1, x_2, ..., x_m)$ follows a multinomial distribution Mult$_{\pi_1, \pi_2, ..., \pi_m}(x_1, x_2, ..., x_m)$ with $\sum_i \pi_i = 1$, where $\pi_i$ is the probability that plot $i$ is picked by an observer, which we can estimate as $\widehat{\pi_i} = x_i/n$.
